---
title: "Cross_validation"
author: "Pei Yang Hsieh"
date: 2018-11-13
output: html_document
---

```{r setup}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Examples

CV "by hand"

```{r}
nonlin_df = tibble(
  id = 1:100,
  x = runif(100, 0, 1),
  y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
)

ggplot(nonlin_df, aes(x = x, y = y)) + geom_point() + theme_bw()
```


Partition into training and testing

```{r}
#training data set 
train_df = sample_n(nonlin_df, 80)

#test data set
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```

Fit a few models

```{r}
#linear model
lin_mod = lm(y ~ x, data = train_df)

#fit smooth curve
nonlin_mod = mgcv::gam(y ~ s(x), data = train_df)

#jagged model
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

Plot the models

```{r}
#this is good
train_df %>% 
  add_predictions(nonlin_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```



```{r}
#this is less good, much too complex
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```


```{r}
train_df %>% 
  gather_predictions(lin_mod, nonlin_mod, wiggly_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

Compute root mean squared errors (RMSEs)

```{r}
rmse(lin_mod, test_df)
## [1] 0.7163422
rmse(nonlin_mod, test_df)
## [1] 0.2437012
rmse(wiggly_mod, test_df)
## [1] 0.3471883
```

CV using modelr

```{r}
#this iterates on number of splits. in this case, does 100 random splits
#gives new data frame with nested list columns
cv_df = 
  crossv_mc(nonlin_df, 100) 
```


Convert to data frame

```{r}
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble

cv_df %>% pull(test) %>% .[[1]] %>% as_tibble

cv_df =
  cv_df %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

Iterate using map, fits model many times

```{r}
#mutate double means it shows up as a number
cv_df = 
  cv_df %>% 
  mutate(lin_mod    = map(train, ~lm(y ~ x, data = .x)),
         nonlin_mod = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
         wiggly_mod = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

plot below shows the distribution of RMSE values for each candidate model. Nonlinear doing best because lowest errors (y-values)

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```


Nepalese children example

```{r}
child_growth = read_csv("./nepalese_children.csv")

child_growth %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5)
```


```{r}
#define pwl split/kink around 7
child_growth =
  child_growth %>% 
  mutate(weight_sp = (weight > 7) * (weight - 7))
```


```{r}
lin_mod = lm(armc ~ weight, data = child_growth)
#piecewise linear model
pwl_mod = lm(armc ~ weight + weight_sp, data = child_growth)
nonlin_mod = gam(armc ~ s(weight), data = child_growth)
```

Plot models

```{r}
child_growth %>% 
  gather_predictions(lin_mod, pwl_mod, nonlin_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

Checck prediction errors for these models

```{r}
cv_df =
  crossv_mc(child_growth, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```


```{r}
cv_df = 
  cv_df %>% 
  mutate(lin_mod = map(train, ~lm(armc ~ weight, data = .x)),
         pwl_mod = map(train, ~lm(armc ~ weight + weight_sp, data = .x)),
         nonlin_mod = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_pwl = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)))
```


Plot prediction error distributions

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```


Piecewise linear fit is what Jeff would go with. It is comparable to the nonlinear and is easier to interpret because only 2 straight lines (not curves).
